\documentclass{article}[11pt,subeqn]

\title{The Twin Instrument\footnote{We are grateful to Paul Devereux, James Fenske, 
Atheen Venkataramani and Cheti Nicoletti,  [WHO ELSE TO ADD?] along with seminar 
audiences and discussants at CMPO Bristol, 
ESPE, NEUDC, CSAE, and The University of Essex for helpful comments.  We make full data and
source code available for examination and replication.  This is available at
\texttt{https://bitbucket.org/damiancclarke/twins}.}}
\author{Sonia Bhalotra\thanks{The University of Essex.  Contact: srbhal@essex.ac.uk} 
\and Damian Clarke\thanks{The University of Oxford.  Contact: damian.clarke@economics.ox.ac.uk}}
\date{\today}

\usepackage[capposition=top]{floatrow}
\usepackage{url}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{dcolumn}
\usepackage{color}
\pagecolor{white}
\usepackage{pdfpages}
\usepackage{lastpage}
\usepackage{lscape}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{breqn}
\usepackage{appendix}
\usepackage{natbib}
\bibliographystyle{abbrvnat}
\bibpunct{(}{)}{;}{a}{,}{,}

\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{multirow}

\usepackage{wrapfig}
\usepackage{blindtext}

\setlength\topmargin{-0.375in}
\setlength\textheight{8.8in}
\setlength\textwidth{5.8in}
\setlength\oddsidemargin{0.4in}
\setlength\evensidemargin{-0.5in}
\setlength\parindent{0.25in}
\setlength\parskip{0.25in}

\newcommand{\person}{we\ }
\newcommand{\Person}{We\ }

\bibliographystyle{abbrvnat}
\bibpunct{(}{)}{;}{a}{,}{,}

\newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}

\usepackage{changepage}% http://ctan.org/pkg/changepage
\makeatletter
\newenvironment{chapabstract}{%
    \begin{center}%
      \bfseries Abstract
    \end{center}}%
   {\par}
\makeatother


\newcommand{\twinfolder}{./../../../Twins}
\newcommand{\biblioinc}{\newpage \bibliography{./../../Document/ThesisRefs}}

\begin{document}
\begin{spacing}{1.4}

\maketitle
\begin{abstract}
 The incidence of twins has been used to identify the impact of changes in fertility 
 on measures of investment in children born prior to the twins, and the emerging 
 consensus in this literature is that there is no evidence of a quantity-quality 
 trade-off. We argue that the standard approach is flawed for two reasons. First, 
 even if twin conception is random, bringing twins to term is a function of maternal 
 health which is difficult to fully observe and which tends to be correlated with 
 child quality, rendering the instrument invalid. Second, twins will only constitute 
 a shock to family size if their occurrence takes family size across the desired 
 level. The neglect of both of these considerations in the literature will tend to 
 lead to under-estimation of the quality-quantity (Q-Q) trade-off and so could 
 contribute to explaining the negative results in the literature. Using a large sample
 of microdata from developing countries, we show that a significant trade-off emerges
 upon correcting for these biases.  We show that this result is likely to be only a 
 \emph{lower} bound of the true Q-Q trade-off and discuss how to estimate the size of
 these bounds. \\
\end{abstract}

\hspace{4mm}\textbf{\small JEL codes}: J13,J16,J18,I15,O15. \\
\newpage
\section{Introduction}
\input{./Introduction.tex}

\section{The Twin Literature}
\input{./TwinLiterature.tex}

\section{Data}
\input{./Data.tex}

\section{Methodology}
\input{./Methodology.tex}

\section{Results}
\input{./Results.tex}

\section{Conclusion}
\input{./Conclusion.tex}




\newpage
\section*{Figures}
\input{\twinfolder/Figures/Figures.tex}
\clearpage

\section*{Tables}
\input{\twinfolder/Tables/Tables.tex}
\clearpage
\newpage

\bibliography{./BiBBase1}

\newpage
\appendix
\section*{Appendices}

\section{Appendix Tables}
\input{\twinfolder/Tables/Appendix_Tables.tex}


\end{spacing}
\end{document}
