\documentclass{nature}
\bibliographystyle{naturemag}
\usepackage{graphicx}
\usepackage[capposition=bottom]{floatrow}
\usepackage{subfloat}
\usepackage{subcaption}
\usepackage{lineno}
\usepackage{lscape}
\usepackage{bm}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{colortbl}

\definecolor{LightCyan}{rgb}{0.7,0.9,1}
\renewcommand\thetable{S\arabic{table}}  
\renewcommand\thefigure{S\arabic{figure}}  
\renewcommand{\thepage}{S\arabic{page}}

\title{Supplementary Information}

\begin{document}

\maketitle

\begin{linenumbers}

The supplementary information for ``Twin Births and Maternal Condition'' provides additional details regarding data and estimation methodologies to support the findings in the paper that the likelihood of a twin birth depends positively upon maternal health stocks and health-related behaviours during pregnancy. A full list of DHS survey data and timing is presented in table S1, while tables S2 and S3 provide conditional regression results to accompany table 1 and figure 3 (respectively) of the paper.  Table S4 documents the effect of smoking during pregnancy on birth weight (a frequently used measure of child health at birth) rather than on twins, and table S5 replicates table 1 panel A, however using \emph{only} the sample of births in which Artificial Reproductive Technologies were used.  Section 2 provides Supplementary Data information on all sources of data.



\clearpage
\section{Supplmentary Tables}
\begin{spacing}{1}
\input{Countries.tex}
\input{twinEffectsCond.tex} 
\begin{landscape}
\input{FDeath_Cond.tex}
\end{landscape}

\begin{table}
  \begin{center}
    \begin{tabular}{lccc}
      \toprule
      \textbf{Dependent Variable:}      & All & Non-Twin & Twin \\
      Birthweight      & Births & Births & Births \\
      \midrule
      Smokes 3 Months Prior to Pregnancy & -98.36*** & -100.8*** & -57.41*** \\
      & (1.172) & (1.189) &  (5.591) \\
      Smokes Trimester 1 & -140.3*** & -144.1*** & -92.67*** \\
      & (1.316) & (1.333) & (6.440) \\
      Smokes Trimester 2 & -163.0*** & -167.8*** & -106.9*** \\
      & (1.390) & (1.407) & (6.940) \\
      Smokes Trimester 3 & -168.3*** & -173.2*** & -109.8*** \\
      & (1.417) & (1.434) & (7.137) \\
      \midrule
      Average Birthweight & 3,283.5   & 3,311.5   & 2,369.7 \\
      Observations        & 1,411,556 & 1,370,368 & 40,151  \\
      \bottomrule
      \caption{\textbf{Smoking and birthweight} {\footnotesize Each cell represents a multivariate OLS regression of smoking behaviour on birthweight using the sample of USA birth data used in table 1.  All specifications fo\
          llow those reported in table 1.  Smoking in each period is a binary measure, and birthweight is measured in grams.}}
    \end{tabular}
  \end{center}
\end{table}
\input{twinEffectsIVF.tex}



\end{spacing}

%%%\begin{table}
%%%  \begin{center}
%%%    \caption{Total Factor Analysis of Health and Twinning}
%%%    \begin{tabular}{lcccc}
%%%      \toprule
%%%\textbf{Dependent Variable:}       &Developed & United & Chile & United \\
%%% Predicted Factor (Z-score)        &Countries & States &       & Kingdom \\ \midrule
%%%      Twin Mother&0.163***&0.0250***&0.111***&0.110*\\
%%%      &(0.0048)&(0.0051)&(0.0343)&(0.0645)\\
%%%      Constant&-0.0008&-0.0822&-0.368&-0.262 \\
%%%  &(0.0007)&(0.0008)&(0.624)&(0.994) \\
%%%      &&&&\\
%%%  Observations&2,053,144&1,363,558&26,581&10,365 \\ \bottomrule
%%%    \end{tabular}
%%%  \end{center}
%%%\end{table}
%%%
%%%
%%%\begin{table}
%%%  \begin{center}
%%%    \caption{Underlying Factor Loadings}
%%%    \begin{tabular}{lcccc}
%%%      \toprule
%%%      &     Factor 1&     Factor 2&     Factor 3&  Uniqueness\\
%%%      \midrule
%%%      \textsc{Developed Countries (DHS)} &&&\\
%%%      Mother's Height (cm)&    .0253809&   -.1675533&    -.111682&    .9588088\\
%%%      Not Obese             &   -.0721951&   -.1984215&     .362284&    .8241671\\
%%%      Not Underweight           &    .0700409&    .0340981&   -.3224643&    .8899483\\
%%%      Attended Births in Area (\% Doctor)&    .3952073&    .8354851&    .0165493&    .1454991\\
%%%      Attended Births in Area (\% Nurse)&     .388454&    -.661938&    .0089997&    .4108599\\
%%%      Attended Births in Area (\% Any)&    .9768874&   -.0694478&    .0002767&    .0408901\\
%%%      \midrule
%%%      \textsc{United States} &&&\\
%%%      Mother's height (cm)&   -.0340336&   -.0315658&    .0309955&    .9968846\\
%%%      Mother Didn't Smoke Before Pregnancy&    .8416049&    .5059564&   -.0007798&     .035777\\
%%%      Mother Didn't Smoke in Trimester 1&    .9270037&    .1197551&    .0004798&    .1263201\\
%%%      Mother Didn't Smoke in Trimester 2&     .988942&   -.1122964&    .0001707&    .0093789\\
%%%      Mother Didn't Smoke in Trimester 3&    .9479534&    -.092264&   -.0009937&    .0928711\\
%%%      Mother Didn't have pre-pregnancy diabetes&    .0057924&    .0038151&    .1580478&    .9749741\\
%%%      Mother Didn't have pre-pregnancy hypertension&     .007462&      .00469&    .2115058&    .9551902\\
%%%      Mother was not obese (pre-pregnancy)&    .0244563&    .0189005&    .4860942&     .762738\\
%%%      Mother was not underweight (pre-pregnancy)&     .041326&    .0044408&   -.2319785&    .9444594\\
%%%      \midrule
%%%      \textsc{Chile} &&&\\
%%%      Did Not Smoke During Pregnancy      &    .2628755&    .1703617&    .0241949&     .901288\\
%%%      Did Not Take Drugs (Moderate) &    .0705569&    .2199089&   -.0217474&    .9461889\\
%%%      Did Not Take Drugs (High)  &    .3689988&   -.1262455&    .0026437&     .847895\\
%%%      Did Not Drink Alcohol (Moderate) &    .1393594&    .2894535&    .0099134&    .8966973\\
%%%      Did Not Drink Alcohol (High) &    .3651861&   -.1491415&    .0184134&    .8440569\\
%%%      Not Low Weight Before Pregnancy    &    .1082982&    .0038736&    -.127083&    .9721064\\
%%%      Not Obese Before Pregnancy        &   -.0009626&    .0090933&    .1499253&    .9774388\\
%%%      \midrule
%%%      \textsc{United Kingdom} &&&\\
%%%      Not Low Weight Before Pregnancy     &    .0086639&    .0639742&    .1024191&    .9853426\\
%%%      Not Obese Before Pregnancy           &   -.0153507&    .0390087&    -.059867&    .9946586\\
%%%      No Hypertension Before Pregnancy    &    .0035539&   -.0110148&   -.0003171&    .9998659\\
%%%      No Diabetes Before Pregnancy    &   -.0174173&   -.0014923&    .0264027&    .9989973\\
%%%      Did Not Drink Alcohol (Moderate)  &    .6167461&   -.1001744&   -.0474115&    .6073414\\
%%%      Did Not Drink Alcohol (High) &     .624899&   -.0191377&   -.0316698&     .608132\\
%%%      No Passive Smoke &    .0733315&    .5084012&     -.04006&    .7345459\\
%%%      Did Not Smoke  &    .0634334&    .5291988&    .0053393&    .7158963\\
%%%      Height              &   -.0567498&    .1121194&   -.0612136&    .9804616\\
%%%      Frequently Consumed Health Food  &   -.0032402&    .1282038&     -.13291&    .9658882\\
%%%      Frequently Consumed Fresh Fruit  &   -.0160699&    .2642819&   -.1342545&    .9118725\\
%%%      
%%%
%%%      \bottomrule
%%%      
%%%
%%%    \end{tabular}
%%%  \end{center}
%%%\end{table}

\section{Supplementary Data}
\subsection{United States National Vital Statistics System (NVSS)}
The NVSS is maintained by the Center for Disease Control and Prevention (CDC). %and collects a range of vital statistics with complete coverage nationwide in the United States.
For each registered birth it makes available the information recorded on the U.S. Standard Certificate of Live Birth. This includes the child's sex, birth type, and indicators of maternal health and health-related behaviors before and during pregnancy. These data have been collected with 100\% coverage from 1972 onwards. We use singleton and twin births from 2009-2013\cite{Martinetal2013} occurring to all mothers aged 18-49 at the time of birth, since 2009 was the first year in which the data record whether or not the mother engaged in artificial reproductive technologies (ART).% an important indicator which may confound our reported results.  Our final estimation sample thus consists of all singleton and twin births occurring in the United States between 2009-2013 
We restrict the sample to births for which ART was not used and drop multiple births involving more than two children (0.09\% of all births). The final estimation sample consists of 13,962,330 births, of which 399,777 (or 2.86\%) are twins.
%2,114,467 (62,340) 2.86
%2,395,397 (70,111) 2.84  
%2,881,080 (84,911) 2.86 
%3,050,474 (89,239) 2.84 
%3,121,135 (93,176) 2.90
  
%did we use all of these below?

%The variables of interest are the child's twin status, and all available measures of maternal health preceding the birth of the child for health behaviours, and preceding the conception of the child for health conditions.  %This includes maternal health behaviours such as smoking during pregnancy, and maternal health stocks such as diabetes and hypertension prior to pregnancy.  All of these variables are directly reported on birth certificates, and available in data. 
%Relevant health measures available from the long form birth certificates are:\ whether the mother smokes in each trimester of pregnancy and in the three months leading up to the pregnancy, whether she suffered from diabetes and hypertension before pregnancy, as well as her height in centimetres, her pre-pregnancy BMI, and total years of education. 
%% I removed above because it is clear in main paper
 %Damian: Can we be precise below ----In certain...do you mean in teh US and Sweden samples?
 %%%%% Sonia: Have updated to say that ``certain'' just refers to all the conditional regressions.
The available indicators are displayed in Table 1. In all conditional regression specifications we also use the reported mother's age, total number of prior births, and the gestational length of the birth as control variables\cite{Hall2003}.  %Summary statistics for these variables are presented in panel A of extended data table 1.
  

%%DC - for later- if Sweden has still births might it have miscarriage so we can study fetal death maybe in our econ paper?
%%%%NOTE: I will check in with Hanna!
%%DC above and below i have added mention of "`Table 1"' - i checked in main paper that there is no latek-code for table 1 it is just table 1 so i hope thats ok.
%%%%Thanks.  That's right.
\subsection{Swedish Medical Birth Register}
%Damian: can we be more precise , why nearly the universse? which are the ``certain specifications''
 %%%%% Sonia: I have added in a line.  The documentation isn't so clear -- just a small portion missing because data isn't perfect... 
The Swedish Medical Birth Register contains data on nearly the entire universe of births occurring in Sweden starting in 1973\cite{EPC2003}.  Approximately 1.4\% of births are not included where completed records are not sent by the delivery hospitals\cite{EPC2003}. The registers record live and still births, the health of the mother and child and birth type (singleton or twin).  Our estimation sample consists of all live singleton or twin births for which full information is available occurring between 1993 and 2012 (a period during which a number of key variables such as smoking are recorded) to all mothers aged 18-49 at the time of birth.  The estimation sample consists of 1,233,546 births, of which 29,482 are twins (2.39\%). 

The maternal health indicators are displayed in Table 1. Smoking is recorded at gestation weeks 10-12 and 30-32, weight is pre-pregnancy weight (from which BMI and indicators of underweight and obesity are calculated) and medical diagnoses are for chronic pre-pregnancy conditions (diabetes, asthma, hypertension, and kidney disease). We observe maternal age, total number of prior births, and gestational length, and include these as control variables in all conditional regression specifications.

\subsection{Avon Longitudinal Study of Parents and Children (ALSPAC)}
ALSPAC is a longitudinal survey that has tracked families since 1990. Our sample consists of the 10,463 ALSPAC children whose mothers responded to health surveys during pregnancy, among whom are 248 twins (2.37\%). The pre-pregnancy maternal health surveys ask a series of questions regarding the mother's diet, physical health and health behaviours. ALSPAC is available to researchers but requires application. The variables we use are in Table 1 where healthy diet refers to the reported frequency of consuming fresh fruit and ``healthy'' foods, alcohol consumption is during pregnancy and the medical conditions are pre-pregnancy. In conditional regression models we control for fixed effects for mother's age and total fertility.
%Damian: for consistency versus sweden? 
% Summary statistics for all variables are presented in panel C of extended data table 1.

\subsection{Chilean Survey of Early Infancy}

%Damian: below- carers or mothers? carers- their births?
 %%%%% Sonia: There are a very small portion of interviewees who are not biological mothers.  In the final sentence of the paragraph I say that we restrict to only biological mothers.
The Longitudinal Survey of Early Infancy is an ongoing nationally representative survey in Chile that started in 2009.  We retain 14,050 of the 15,175 cases which consist of biological mothers with complete records for children. In total, 358 twin births (2.55\%) are observed. Maternal health indicators are as in Table 1, where pre-pregnancy weight is used to construct indicators of  `normal', `underweight' and `obese', smoking, alcohol and drug consumption are during pregnancy and the survey asks if these behaviours were never, sporadic, or regular. The omitted categoy is ``never consumed''. We also have mother's age, region of residence, and total completed fertility as conditioning variables. 

\subsection{Demographic and Health Surveys}
The Demographic and Health Surveys (DHS), funded by USAID, are nationally representative and have a module focusing on fertile aged women, which records measures of health, fertility, and birth type (twin/singleton).  We collect all publicly available surveys conducted between 1990 and 2013, resulting in data from 68 countries on 2,058,324 singleton or twin births, of which 43,020 (2.09\%) are twins.  Births with multiplicity of greater than 2 (which constitute less than 0.008\% of the sample) are removed.  A full list of the surveys included in our sample is available in Supplementary Information table S1.

%Our estimation sample consists of all births in DHS data occurring to women aged between 18 and 49 years, for whom height and BMI were recorded.  We remove from the sample any women with a recorded height greater than 240cm or less than 70cm, and those with a BMI greater than 50, as these are likely to  be recording errors.  This results in a sample of children born between 1961 and 2012.  

% Damian i know it is too late to get it in now but just between us do you know what fraction of countries record anemia i thought many did, i used in my work with S Rawlings
%%%%% Sonia: I looked into this, and the main problem isn't coverage (which as you say isn't so bad -- at least from memory), but timing of the measure.  Anemia is measured at the time of survey and twins come from the fertility history, so in all cases the anemia measure is *after* the twins are born, which is problematic.  

%%DC i added discussion of the BMI bias in DHS below- see if you think we should keep it or just leave it out. It's all less important if it is in supplementary material as i suggest above
%%%%This is useful, thanks for adding.  Also now in supplementary material it's not a problem to have further details.
In the DHS, BMI is recorded at the time of the survey rather than pre-pregnancy and births to women who are 18 to 49 years old at the survey date occur anywhere between the survey date and some 30 years before the survey date, but as long as distance from survey is balanced between twin and non-twin mothers, it seems valid to use BMI at survey as a proxy for long-term BMI. The DHS contains indicators of medical facilities in the woman's survey cluster. Clusters  are the primary sampling units and, in our sample, the average number of children in a cluster is 87. In order to determine the availability of medical care (rather than actual usage which would contain selection), we calculate the proportion of all pregnancies in a cluster which were attended by doctors, attended by nurses, attended by village health workers, and unattended by any medical professionals during the prenatal period. These variables are constant at the level of the cluster, and in regressions, prenatal care from a village health workers is used as the omitted base category. Since we pool DHS data across countries and survey years, we consistently include country and year fixed effects. %Summary statistics for all variables are presented in panel E of extended data table 1.

%%DC Sorry i did not raise this earlier but editing the above line it occured to me that when we add controls for ma age and b-order we should really be adding controls for child birth year. One might even argue we should add controls for birth year in the unconditional regs since the different samples are for different birth years and we should be allowing for trends in both twinnnig and maternal health indicators. what do you think? 
%Damian: can we define the 3rd var in T1 here? 
%%%%FORTUNATELY WE HAVE THESE (The only case where it really matters is DHS, as the rest are shorter time ranges (as few as two years).  We are indicating this as \phi_t in regression equation (1) and (2), and also in (5). 
%%%%% Sonia: Just to check, do you refer to the third variable in the summary statistics table?
%%DC- i referred to Table 1 "`prenatal care availability"' which comes after doctor and nurse. above we define only doct andor and nurse. we need to define prenatal availability. is it existence of a clinic or non-zero uptake or something else.
%%DC more important in T1 given the dr, nurse defs above i think we should replace "`Doctor Availability"' with "`Doctor attended birth"' and same for Nurse assuming that these are the vars available.
%Damian above you mention the categories but what about unattended births? 
%%%% DONE: This was my mistake!  Now I see where the confusion was coming from.  I meant to write that these were always prental care measures (% with prenatal care), and not percent attended births.  We have written this correctly in the body of the paper (and also in all tables except for the summary statistics), so the mistake was here.  I have now changed it to make it clear that we mean PRENATAL care, and not attended births.  I think this is now all consistent, and it is hopefully much clearer what the three variables are.
%%%%% Sonia: Fixed.  Earlier it was "attended by any", now I have said "unattended by any".


\subsection{Foetal Death Data}
In order to compare the likelihood of foetal death for twins and singletons, we combine foetal death and birth data from the National Vital Statistics Data. States are required to report foetal deaths to the National Vital Statistics System when they occur at greater than 20 weeks of gestation, or weighing more than 350 g. A number of states report deaths occurring at less than 20 weeks, however for comparability we restrict our sample to foetal deaths occurring at or after 20 weeks of gestation.  We focus on foetal deaths occurring in the years prior to 2003.  In recent years (2008 onwards) foetal death data do not contain information on maternal health, and between 2003 and 2008 fewer measures of health were included.  Prior to the 2003 revision of the birth certificate however, identical maternal health questions were recorded in the foetal death and live birth records.
%Damian What about 2003-2008?
%%%%% Sonia: Added this message above.  Basically, pre 2003 is the best year for full coverage of all health variables in both foetal death and live birth data.  While it would be possible to do *something* between 2003-2008 it would require some cautious coding, and various states would drop out, as from 2003 onwards states changed over to the new birth certificate format in different years.

Our sample consists of all singleton or twin live births and foetal deaths (occurring at 20 weeks of gestation or later) recorded in the NVSS during 1999-2002, which results in a sample of 13,411,182 pregnancies, 66,251 of which resulted in foetal deaths. Comparable maternal health variables in both birth and death data files are those displayed in Figure 3 and Extended Data Table E3. %(%%DC: xx is the table with the regression results for miscarriage which i feel should be in extended data but i dont feel v strongly if you want to leave it in Supplem for now). 
Summary statistics are in extended data table E4.

\clearpage
\bibliography{refs}

%\end{spacing}
\end{linenumbers}
\end{document}
