\documentclass{article}[11pt,subeqn]

\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage[pdftex]{graphicx}
\usepackage{setspace}
\usepackage{framed}
\usepackage{lastpage}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\title{Child Quantity versus Quality: Are Twin Births Exogenous?}
\author{Damian C. Clarke}

\setlength\topmargin{-0.375in}
\setlength\headheight{20pt}
\setlength\headwidth{6.5in}
\setlength\textheight{8.7in}
\setlength\textwidth{6.5in}
\setlength\oddsidemargin{0in}
\setlength\evensidemargin{-0.5in}
\setlength\parindent{0.25in}
\setlength\parskip{0.25in}

\usepackage{natbib}
\bibliographystyle{abbrvnat}
\bibpunct{(}{)}{;}{a}{,}{,}

\usepackage{lscape}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{rotating,capt-of}
\usepackage{array}

\usepackage{lineno}

%NEW COMMANDS
\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\vect}[1]{\mathbf{#1}}
\newcolumntype{P}[1]{>{\raggedright}p{#1\linewidth}}

\usepackage{appendix}
\usepackage{booktabs}
\usepackage{cleveref}

\fancyhead{}
\fancyfoot{}
\fancyhead[L]{Extended Essay}
\fancyhead[C]{MSc. Economics for Development}
\fancyhead[R]{Damian C. Clarke}
\fancyfoot[C]{\thepage\ of  \pageref{LastPage}}
\fancyfoot[R]{DRAFT}

\begin{document}
\linenumbers
\begin{spacing}{1.25}

\maketitle
\begin{abstract}
The justification of family planning policies in low income countries may draw some theoretical support 
if the production of child quality by a household is a function of the number siblings also competing
for investment.  This hypothesis has been tested using twins as an exogenous shock to family size.  Here, I test
the exogeneity of twins to family size, and derive the bias that may occur if twinning is actually
an endogenous event.  
\end{abstract}

\section{Introduction}
\citet{Malthus1798} famously introduces a theory in which population growth enters into competition with economic   
well-being. Contemporary microeconomic models have described the existence of similar mechanisms at a household level, albeit with less disastrous consequences.  Early work by \citet{BeckerLewis1973} and \citet{BeckerTomes1976} propose that a household's ability to produce child `quality' is a function of the quantity of children.  The evidence in favour of this quantity-quality (Q-Q) relationship is significant.  Surveys at both a macro- and microeconomic level have documented the existence of a negative correlation between achievement measures and family size (see for example \cite{Desai1995}, or \cite{Hanushek1992}).

TAKE OUT THESE TWO PARAGRAPHS (THEY ADD LITTLE AND AT BEST SHOULD BE IN LIT REVIEW).  REPLACE WITH A DISCUSSION OF WHY Q-Q TRADEOFF MAY EXIST AND WHY THIS IS PARTICULARLY SO IN THE DEVELOPED WORLD.
At the household level, the Q-Q model is supported by the frequently observed empirical relationship between an individual's sibship size and a
range of measures of their `quality' such as completed years of schooling, mortality, or labour force participation.
The estimation of the causal relationship between quality and quantity is less clear. 

Recent empirical results have called into question the validity of the Q-Q trade-off hypothesis.  The majority of
these studies however have been based upon data in developed countries, where often educational investments
required by parents are highly subsidised by the State.  It seems important to test this hypothesis in developing
countries, where the production function for child quality may be substantially different, requiring more intensive
investments by the family.  A number of recent papers have used Chinese data and found results
more in line with the Q-Q model.  Here, I extend this analysis to a wider range of developing countries, providing
a novel examination relevant to populations facing a stricter (average) household budget constraint.
\vspace{-5mm}
\section{Literature}
\label{scn:lit}
\vspace{-5mm}
\input{LitReview.tex}
\section{Data}
\label{scn:data}
\vspace{-5mm}
In order to address this question, the dataset used must include various measures of both mother and child
outcomes. In order to estimate specification (\ref{eqn:twin}), information must be available on a mother's
health and socioeconomic outcomes along with details of parity -- principally birth order, the presence
of multiple births, and the gender of each child.  It is important that this data be representative at a
national level.  This regression will be estimated using data from the Demographic and Health Surveys (DHS),
a nationally representative survey currently administered in approximately 90 developing countries.

In order to estimate the Q-Q model (\ref{eqn:model}), additional data requirements exist.  Principally,
information must be available regarding `quality' outcomes for children, as well as regarding the mother's
health and economic status.  This requires that children are observed late enough in life to have suitable
quality measures.  This data will be obtained from Chile's Social Protection Survey (EPS for its initials
is Spanish), a nationally representative panel survey, which follows all members of a household.  The panel
nature of this survey will allow for the inclusion of lagged health and socioeconomic characteristics,
allowing for the mother's characteristics earlier in her children's lives to measured.  In order to test
the power of these findings in light of earlier empirical work, I will recreate past analyses with the
data used (such as Caceres, 2006), giving some indication of the degree of importance of any biases observed.

\begin{table}[ht]
\begin{center}
\begin{tabular}{lcc} \toprule
 VARIABLES & Height Recorded & Missing \\ \midrule
Twin & 0.0240 & 0.0202 \\
\begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.153)\end{footnotesize} & \begin{footnotesize}(0.141)\end{footnotesize} \\
Fertility &  5.36 & 5.45\\
\begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(2.81)\end{footnotesize} & \begin{footnotesize}(2.83)\end{footnotesize} \\
Educfyrs & 3.40 & 3.58\\
\begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(4.18)\end{footnotesize} & \begin{footnotesize}(4.11)\end{footnotesize} \\
Contraceptive\_use &  1.69 &  1.59 \\
\begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(1.42)\end{footnotesize} & \begin{footnotesize}(1.37)\end{footnotesize} \\
Contraceptive\_know &  2.79 & 2.54 \\
\begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.757)\end{footnotesize} & \begin{footnotesize}(0.900)\end{footnotesize} \\
Poor1 & 0.0882 & 0.105\\
\begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.284)\end{footnotesize} & \begin{footnotesize}(0.307)\end{footnotesize} \\ \midrule
Observations & 2,139,076 &  2,093,323  \\ \bottomrule
\end{tabular}
\caption{Summary statistics of missing versus non-missing height}
\label{tab:missing}
\end{center}
\end{table}


\section{Empirical Framework}
A 2SLS estimation strategy to estimate the causal effect $\alpha$ in (\ref{eqn:model}) involves the following
regressions:
\begin{subequations}
\label{eqn:2SLS}
\begin{eqnarray}
Q_{ij}&=&\beta_0+\beta_1Size_j+\beta_2B_{j}+\beta_3X_{ij}+\beta_4W_j+u_{ij} \label{eqn:2SLSa}\\
Size_{ij}&=&\gamma_0+\gamma_1Twin_{ij}+\gamma_2B_{j}+\gamma_3X_{ij}+\gamma_4W_j+v_{ij} \label{eqn:2SLSb}
 \end{eqnarray}
\end{subequations}
where $Size_j$ refers to number of children in the family, $B_j$ birth order of the child, $X_{ij}$ a vector
of individual characteristics and $W_j$ a vector of family characteristics such as income and parents' education.
In what follows, exogenous variables in equation \ref{eqn:2SLSa} will be denoted $\mathbf{x_2}$, the
endogenous variable $Size_j$ as $x_1$, the instrumental variable twin as $z_1$, and the outcome variable
$Q_{ij}$ as $y$.  Then, the regressors from equation \ref{eqn:2SLSa} are represented as
$\vect{x}=[x_1\ \vect{x}_2']'$ and the instruments from \ref{eqn:2SLSb} as $\vect{z}=[z_1\ \vect{x}_2']'$.

In order to derive the consistency of the estimators $\vect{\beta}$, we start from the typical instrumental variables estimator:
\begin{eqnarray}
\label{eqn:IVconsist}
\vect{\beta}_{IV}=(\vect{Z}'\vect{X})^{-1}\vect{Z}'\vect{y} \nonumber
\end{eqnarray}
where $\vect{Z}$ is an $N \times 4$ matrix with $i$\textsuperscript{th} row $\vect{z}_i'$.  Determining consistency proceeds via
the susbstitution of the structural equation for $\vect{y}$:
\begin{eqnarray}
\label{eqn:IVconsist2}
\vect{\beta}_{IV}&=&(\vect{Z}'\vect{X})^{-1}\vect{Z}'[\vect{X\beta}+\vect{u}] \nonumber\\
&=&\vect{\beta}+(\vect{Z}'\vect{X})^{-1}\vect{Z}'\vect{u}\nonumber\\
&=&\vect{\beta}+(N^{-1}\vect{Z}'\vect{X})^{-1}N^{-1}\vect{Z}'\vect{u}\nonumber
\end{eqnarray}
where the final line is included in order to demonstrate consistency via the use of the law of large numbers\footnote{This relies
on the fact that $N^{-1}\vect{Z}'\vect{X}=N^{-1}\sum_i\vect{x}_i\vect{x}_i'$ and likewise for $N^{-1}\vect{Z}'\vect{u}$.  A
complete exposition is provided in Cameron and Trivedi (2006).}.  Consistency of the IV estimator requires that
\begin{eqnarray}
\label{eqn:IVconsist3}
\mathrm{plim} N^{-1}\vect{Z'u}&=&0, \hspace{5mm}\mathrm{and} \\
\mathrm{plim} N^{-1}\vect{Z'X}&\neq&0,
\end{eqnarray}
a set of conditions very similar to those outlined in equation \ref{eqn:IV}.

This paper is concerned with the potential endogeneity of twin births.  Where twin births are endogenous, responding to
characteristics such as mother's health and education and family income, omitting these factors in the structural equation
\ref{eqn:2SLSa} will lead to biased and inconsistent estimates of $\vect{\beta}_{IV}$.

\section{Results}
\label{sec:results}
\subsection{Twin Endogeneity}
\label{sec:twinendog}
Blah blah blah.  Table \ref{tab:twinreg1} presents regressions for the probability of multiple birth using the full sample of DHS data.
\begin{equation}
\label{eqn:twinpred}
P(t_i=1)=\gamma_1 y_i + \gamma_2 h_i + \gamma_3 m_i +X_i'\beta + \epsilon_i.
\end{equation}

\begin{table}[ht]
\begin{center}
\begin{tabular}{lcc} \toprule
  & (1) & (2) \\
 VARIABLES & twind & twind \\ \midrule
 \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}\end{footnotesize} \\
 fertility & 0.0825*** & 0.00490*** \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.00330)\end{footnotesize} & \begin{footnotesize}(0.000352)\end{footnotesize} \\
  educfyrs & 0.0125*** & 0.000704*** \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.00204)\end{footnotesize} & \begin{footnotesize}(0.000116)\end{footnotesize} \\
  contracept\_use & 0.00660 & 0.000218 \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.00505)\end{footnotesize} & \begin{footnotesize}(0.000275)\end{footnotesize} \\
  contracept\_know & 0.0123* & 0.000688 \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.00727)\end{footnotesize} & \begin{footnotesize}(0.000418)\end{footnotesize} \\
  height & 0.0125*** & 0.000673*** \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.00111)\end{footnotesize} & \begin{footnotesize}(5.13e-05)\end{footnotesize} \\
  bmi & 0.00233 & 9.72e-05 \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.00226)\end{footnotesize} & \begin{footnotesize}(0.000126)\end{footnotesize} \\
  poor1 & -0.117*** & -0.00567*** \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.0394)\end{footnotesize} & \begin{footnotesize}(0.00192)\end{footnotesize} \\
  Constant & -4.559*** & -0.114*** \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}(0.175)\end{footnotesize} & \begin{footnotesize}(0.00773)\end{footnotesize} \\
  \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}\end{footnotesize} & \begin{footnotesize}\end{footnotesize} \\
  Observations & 2,083,707 & 2,083,707 \\
  R-squared &  & 0.009  \\ \midrule
\multicolumn{3}{c}{\setstretch{0.9}\begin{footnotesize}Robust standard errors in parentheses\end{footnotesize}} \\
\multicolumn{3}{c}{\begin{footnotesize}*** p$<$0.01, ** p$<$0.05, * p$<$0.1\end{footnotesize}} \\ 
\multicolumn{3}{p{7.2cm}}{\setstretch{0.9}\begin{footnotesize}\textsc{Note:} Column (1) represents probit with MFX at the mean.  Column
(2) is OLS.\end{footnotesize}}\\
\bottomrule
\end{tabular}
\end{center}
\label{tab:twinreg1}
\caption{Probability of giving birth to multiple children}
\end{table}
\setstretch{1.25}
\subsection{Bias in IV Regressions}
\subsubsection{Monte Carlo Simulation}
\label{scn:MCS}
The results from section \ref{sec:twinendog} suggest that twin birth is not exogenous. As a result, IV estimates failing to control for the full set of 
relevant variables in the second stage will produce inconsistent and biased estimates of the importance of family size on child quality [REHASH THIS SENTENCE.  IT BRINGS UNNECESSARY CHANCE OF TROUBLE].  As a first approach at quantifying the importance of this bias, Monte Carlo simulations are run allowing for correlation between twin birth and previously uncontrolled factors
to vary.\footnote{Whilst it may seem unlikely that the correlation between twin births and these unobserved factors will vary over time and hence that
simulation of a range of values is un-ilustrative, these simulations may be useful in abstracting to other situations.  This analysis is identical to the
case where multiple births occurr more frequently where in-vitro fertilisation (IVF) is used.  As IVF treatment often involves the implantation of 
multiple embryos, multiple births are more likely here compared with traditional methods of fertilisation (cite).  These simulations then could be 
thought as analogous to examining the role increasing use of IVF plays on the consistency of twin estimates where only birth records are available, 
but details regarding fertilisation are unknown.}  This allows for the examination of the importance of the instrumental validity assumption, essentially
by simulating values of equation (\ref{eqn:IVconsist3}) which increasingly diverge from 0. 

I simulate an ($n \times 3)$ vector of $N=100,000$ standard normal error terms are for the following system of equations.\\
\begin{eqnarray}
\label{eqn:MC1}
y_i=\beta_0+\beta_1 x_i + \epsilon_{1i} \nonumber\\
x_i=\gamma_0 + \gamma_1 z_i + \epsilon_{2i} \nonumber\\
z_i=1[\alpha_0+\epsilon_{3i}>0], \nonumber
\end{eqnarray} 
where simulated $\epsilon_j$ follow:
\begin{equation}
\begin{bmatrix}
\epsilon_1\\
\epsilon_2\\
\epsilon_3\\
\end{bmatrix}
\sim \mathcal{N}
\left(\begin{bmatrix}
0\\
0\\
0\\
\end{bmatrix}
,
\begin{bmatrix}
\sigma_{\epsilon_1}^2 &  \rho_{\epsilon_1\epsilon_2} &  \rho_{\epsilon_1\epsilon_3}\\
\rho_{\epsilon_2\epsilon_1} & \sigma_{\epsilon_2}^2 &  \rho_{\epsilon_2\epsilon_3} \\
\rho_{\epsilon_3\epsilon_1}&  \rho_{\epsilon_3\epsilon_2}& \sigma_{\epsilon_3}^2 \\
\end{bmatrix}\right)
\end{equation}


\subsubsection{Empirical Evidence}
\label{scn:EE}
\subsection{Twin Selection}
Sections \ref{scn:MCS} and \ref{scn:EE} demonstrate that the incidence of non-observed twin predictor variables will invalidate the instrumental variable identification strategy.  Monte Carlo simulations suggest that even a relatively minor correlation between these unobserved terms and the stochastic component of the Q-Q equation can result in estimators which perform worse than traditional OLS estimators.  However, if we are able to collect data on these variables which influence twin-selection\footnote{I refer to twin ``selection" here not because families choose whether or not to have twins, but rather because certain characteristics make it appear as if families are more likely to select in to multiple births.  The usefelness of this term will become apparent in the analysis which follows, which draws from prior literature on selection on observables versus selection on unobservables.  Once again, a parallel may be drawn with the use of twins as an instrument in situations where IVF treatment is available; in this case it seems less farfetched to speak of ``selection'' of multiple births.} a 2SLS methodology can still be used consistently by including these predictors as controls in the first and second stage.  The credibility of this strategy however relies on the ability to perfectly observe determinant factors which increase the propensity of twin birth.  If we believe that models such as (\ref{eqn:twinpred}) are structural representations of twin birth, we can simply include all relevant predictors, and recover consistent estimates of family size.

It seems unlikely that survey data will allow for all such relevant predictors to be observed however.  While mother's weight and height and family socio-economic characteristics are observable, many aspects of a mother's pre-natal behaviour are not observed by the econometrician.  Dietary habits and other household investments in (pre-natal) child health may be both difficult to observe, and correlated with child quality and quantity preferences.  In this case, a problem of selection on unobservables exists for twin births.  The inclusion of ovbservable predictors of twin births in the 2SLS specification will not correct for this bias.

 



\section{Conclusion}

\newpage
\bibliography{BiBBase}

\newpage
\appendix
\section{Twin Estimation Strategies}
\label{scn:litrev}

\vspace{19.2cm}	

\begin{center}
\begin{rotate}{90}
\begin{tabular}{lp{4mm}lll}\toprule
Author& &  Estimation Strategy & \multicolumn{2}{c}{Controls} \\ \midrule
&& & \hspace{5mm}1\textsuperscript{st} Stage Controls & 2\textsuperscript{nd} Stage Controls \\  \cmidrule(r){4-4} \cmidrule{5-5}
Rosenzweig and Wolpin (1980) & &
$ED=\omega_0+\omega_1 TR+u_1$&
\begin{small}None.\end{small}&
\begin{small}N\slash A\end{small}.
\\
\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}\\
\multirow{2}{*}{Black \emph{et al.} (2005)} & &
(1) $FSIZE=\alpha_0 + \alpha_1 T + X\alpha_2 + v$ &
\begin{small}Full family size and  birth \end{small}&
\begin{small}Covariates from\end{small}
\\
& &
(2) $ED=\beta_0+\beta_1 FSIZE + X\beta_2 + \epsilon$ &
\begin{small}order dummies. Demogra-\end{small}  &
\begin{small}stage 1.\end{small}
\\
& & &
\begin{small}phic controls: age, sex, pa-\end{small}&
\\
& & &
\begin{small}rental age and education.\end{small}&
\\
\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}\\
\multirow{2}{*}{C\'aceres (2006)} & &
(1) $n_i=X'_i\mu + \rho mb_i + v_i$ &
\begin{small}Dummies by age, state of\end{small}&
\begin{small}Covariates from\end{small}
\\
& &
(2) $y_i=\alpha + \gamma n_i + X'_i\beta + \epsilon_i$ &
\begin{small}residence, mother's educ- \end{small}&
\begin{small}stage 1.\end{small}
\\
& & &
\begin{small}ation, race, mother's age\end{small}&
\\
& & &
\begin{small}and sex.\end{small}&
\\
\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}&\begin{footnotesize}\end{footnotesize}\\
\multirow{2}{*}{Angrist \emph{et al.} (2006)} &  &
(1)  $c_1=X'_i\beta+\alpha t_{2i}+\eta_i$ &
\begin{small}Mother's age, mother's \end{small} & 
\begin{small}Covariates $X_i$\end{small}
\\
& &
(2) $y_i=W'_i\mu+\rho c_i + \epsilon_i$ &  
\begin{small}age at first birth, census \end{small} &
\begin{small}from stage 1.\end{small}
\\
& & &
\begin{small}year, parental birth place.\end{small}&
\\
\bottomrule 
\multicolumn{5}{p{19.2cm}}{\setstretch{0.9}\begin{footnotesize}\textsc{Note:} Nomenclature employed here is as faithful as possible to the original studies.  $TR$ refers to twin ratio, the number of twin births per mother divided by number of pregnancies, and $ED$ represents a child's standardised educational attainment. $FSIZE$ refers to sibship size and $T$ a dummy for twin birth on the $n$\textsuperscript{th} birth, restricting the analysis to children with birth order $<n$. The variable $n_i$ defined by C\'aceres once again refers to sibship size, and $mb_i$ multiple births. Finally, $c_i$ refers to child $i$'s sibship size, with $t_{2i}$ denoting multiple second births.\end{footnotesize}}\\
\end{tabular}
\end{rotate}
\captionof{table}{Prior empirical specifications}
\label{tab:litrev}
\end{center}
\setstretch{1.25}

\section{Full Survey Data}
\label{scn:surveys}
\input{Appendix_SurveyCountries.tex}

\end{spacing}
\end{document}

________________________________________________________________________________________________________________________________



































Naive OLS is
likely to overestimate the importance of the degree of causality.  Where unobserved factors such
as parental preferences drive both quality of children and household size, estimates of the
trade-off will be biased.  Various plausible channels for this mechanism have been proposed, such
as a preference for lower children by more highly educated parents (Qian, 2005). Frequently, data
on multiple births is used to recover a causal estimate of the Q-Q trade-off. Rosenzweig and Wolpin
(1980) discussed the theoretical validity of instrumenting for family size ($N$) with twin birth ($t$),
with this method subsequently receiving considerable treatment in the empirical literature.  The
validity of this estimation strategy relies upon typical IV assumptions, namely
\begin{equation}
\label{eqn:IV}
\begin{aligned}
\text{cov}(t,N)\neq0 \\
\text{cov}(t,u)=0
\end{aligned}
\end{equation}
where u refers to the unobserved error term in the QQ model:
\begin{equation}
\label{eqn:model}
Q_i=\alpha N_i + X_i'\beta + u_i.
\end{equation}

I propose to examine the exogeneity of the instrumental strategy outlined in (\ref{eqn:IV}).
I will ask whether multiple births truly are an exogenous shock to family size, or whether these are
influenced by other socioeconomic factors not considered in prior studies.  Fundamentally, I am
interested in determining whether a mother's likelihood of giving birth to multiple infants depends upon
her household income ($y_i$), investment in own health ($h_i$), or other individual or household
characteristics such as marital status ($m_i$).  These results are quite simple to obtain using a limited
dependent variable regression model (\emph{e.g.}\ a probit) to estimate:
\begin{equation}
\label{eqn:twin}
P(t_i=1)=\gamma_1 y_i + \gamma_2 h_i + \gamma_3 m_i +X_i'\beta + \epsilon_i.
\end{equation}
Prior literature has failed to take account of this threat to instrumental validity, generally assuming
that twin births are exogenous (beyond race and birth-order).

It seems plausible that even in the case that twin conception is exogenous to a household or mother's
socioeconomic characteristics, the likelihood that a mother carries to full term and gives birth to
multiple live infants will not be exogenously determined.  I am interested in testing the null that
$\gamma_1=0$, $\gamma_2=0$ and $\gamma_3=0$, individually and collectively against the alternative
that $\vec{\gamma}>0$.

If this hypothesis is borne out empirically, previous Q-Q model estimates obtained instrumenting for
multiple birth, but not cotrolling for a mother's socioeconomic characteristics, will be biased. In this
essay, the bias will be derived mathematically, and tested empirically.  Using data from Chile (see
section \ref{scn:data}) I will estimate (\ref{eqn:model}) via OLS, IV (not controlling for additional
mother characteristics) and IV controlling for these characteristics.  This will allow me to empirically
determine if the bias in previous estimates exists, is important, and agrees with theoretical
derivations.



\setstretch{1.1}
\begin{eqnarray}
\label{eqn:MC1}
y_i=\beta_0+\beta_1 x_i + \epsilon_{1i} \nonumber\\
x_i=\gamma_0 + \gamma_1 z_i + \epsilon_{2i} \nonumber\\
z_i=1[\alpha_0+\epsilon_{3i}>0], \nonumber
\end{eqnarray} 
where simulated $\epsilon_j$ follow:
\begin{equation}
\left(
\begin{array}{c}
\epsilon_1\\
\epsilon_2\\
\epsilon_3\\
\end{array}
\right)
\sim \mathcal{N}\left(
\Biggl(
\begin{array}{c}
0\\
0\\
0\\
\end{array}
\Biggl),
\Biggl(\begin{array}{c}
\sigma_{\epsilon_1}^2 \quad  \rho_{\epsilon_1\epsilon_2} \quad  \rho_{\epsilon_1\epsilon_3}\\
\rho_{\epsilon_2\epsilon_1} \quad \sigma_{\epsilon_2}^2 \quad  \rho_{\epsilon_2\epsilon_3} \\
\rho_{\epsilon_3\epsilon_1}\quad  \rho_{\epsilon_3\epsilon_2}\quad \sigma_{\epsilon_3}^2 \\
\end{array}\Biggl)
\right)
\end{equation}
\subsubsection{Empirical Evidence}
\label{scn:EE}
\setstretch{1.25}
